Carrier phase and amplitude estimation for phase shift keying using pilots and data

### 2012-12-19 paper.tex
fixed a few estimator - parameter, parameter - estimator things.

In the literature it has been common to assume that the data symbols $\{d_i, i \in D\}$ are of primary interest and that the complex amplitude $a_0$ is a nuisance parameter.  The metric of performance is correspondingly the \emph{symbol error rate}, or \emph{bit error rate}.  While estimating the symbols (or more precisely the transmitted bits) is ultimately the goal, we take the opposite point of view here.  Our aim is to estimate $a_0$, and we treat the unknown data symbols as nuisance parameters.  This is motivated by the fact that in many modern communication systems the data symbols are \emph{coded}.  For this reason raw symbol error rate is not of interest at this stage.  Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated received symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.  The additive noise channel is a common assumption for subsequent receiver operations, such as decoding.  The estimator $\hat{a}$ is also used in the computation of decoder metrics for modern decoders, and for interference cancellation in multiuser systems.  Consequently, our metric of performance will not be symbol or bit error rate, but $\abs{a_0 - \hat{a}}^2$. It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ and $(\rho_0 - \hat{\rho})^2$.  The function $\fracpart{\cdot}_\pi$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.  It will become apparent why $\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ rather than $(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.

@@

The theorems place conditions on $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ rather than directly on $\theta_0 - \hat{\theta}$.  This makes sense because the phases $\theta_0$ and $\theta_0 + 2\pi k$ are equivalent for any integer $k$. So, for example, we expect the phases $0.99\pi$ and $-0.99\pi$ to be close together, the difference between them being $\vert\sfracpart{-0.99\pi - 0.99\pi}_\pi\vert = 0.02\pi$, and not $\vert -0.99\pi - 0.99\pi\vert = 1.98\pi$.

@@

A key assumption in Theorem~\ref{thm:consistency} is that $G(x)$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.  This assumption asserts that $G(x) \leq G(0)$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $G(x_i) \rightarrow G(0)$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.  Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.  One can check that the assumption holds when $w_1$ is circularly symmetric and normally distributed.  We will not attempt to further classify those distributions for which the assumption holds here.

>>>>

In the literature it has been common to assume that the data symbols $\{d_i, i \in D\}$ are of primary interest and that the complex amplitude $a_0$ is a nuisance parameter.  The metric of performance is correspondingly the \emph{symbol error rate}, or \emph{bit error rate}.  While estimating the symbols (or more precisely the transmitted bits) is ultimately the goal, we take the opposite point of view here.  Our aim is to estimate $a_0$, and we treat the unknown data symbols as nuisance parameters.  This is motivated by the fact that in many modern communication systems the data symbols are \emph{coded}.  For this reason raw symbol error rate is not of interest at this stage.  Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated received symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.  The additive noise channel is a common assumption for subsequent receiver operations, such as decoding.  The estimator $\hat{a}$ is also used in the computation of decoder metrics for modern decoders, and for interference cancellation in multiuser systems.  Consequently, our metric of performance will not be symbol or bit error rate, but $\abs{\hat{a} - a_0}^2$. It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\hat{\theta} - \theta_0\rangle_\pi^2$ and $(\hat{\rho} - \rho_0)^2$.  The function $\fracpart{\cdot}_\pi$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.  It will become apparent why $\langle\hat{\theta} - \theta_0\rangle_\pi^2$ rather than $(\hat{\theta} - \theta_0)^2$ is the appropriate measure of error for the phase parameter.

@@

The theorems place conditions on $\sfracpart{\hat{\theta} - \theta_0}_\pi$ rather than directly on $\theta_0  - \theta_0$.  This makes sense because the phases $\theta_0$ and $\theta_0 + 2\pi k$ are equivalent for any integer $k$. So, for example, we expect the phases $0.99\pi$ and $-0.99\pi$ to be close together, the difference between them being $\vert\sfracpart{-0.99\pi - 0.99\pi}_\pi\vert = 0.02\pi$, and not $\vert -0.99\pi - 0.99\pi\vert = 1.98\pi$.

@@

A key assumption in Theorem~\ref{thm:consistency} is that $G(x)$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.  This assumption asserts that $G(x) \leq G(0)$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $G(x_i) \rightarrow G(0)$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.  Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\hat{\theta} - \theta_0}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\hat{\theta} - \theta_0}_\pi$ will not converge.  One can check that the assumption holds when $w_1$ is circularly symmetric and normally distributed.  We will not attempt to further classify those distributions for which the assumption holds here.
